---
title: "HW4 STAT 5014"
author: "Shane Bookhultz"
date: "September 25, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}
library(xlsx)
library(knitr)
library(ggplot2)
library(GGally)
library(dplyr)
```


## Problem 3

The focus of the Exploratory Data Analysis (EDA) stage of analysis is to refine the question/hypothesis, and then from there to know which modeling techniques to use.

## Problem 4 

```{r echo = F}

prob4_data1 <- read.xlsx("HW4_data.xlsx", sheetIndex = 1)

# need to show summary stats
# need to show a factor exploration
# create a couple multipanel plots using either lattice, ggplot2, or baseR
# correlation plots, pairs, plotcorr

ls(prob4_data1)
lapply(prob4_data1, class)
SumProb4Data <- summary(prob4_data1)

prob4_data1$block <- factor(prob4_data1$block)

kable(SumProb4Data)

# showed summary stats, gave a bit of factor explanation

# first create a scatter plot in ggplot2 based on factor groups

plot <- ggplot(prob4_data1, aes(depth, phosphate))

plot + geom_point(aes(colour = factor(block)), size = 1)

# now multipanel plot
attach(prob4_data1)
pairs(depth~phosphate) # R squared is 0.004

hist(depth)
hist(phosphate)
boxplot(depth)
boxplot(phosphate)

ggplot(prob4_data1, aes(x=depth, y=phosphate, fill = block)) + geom_boxplot()

ggplot(prob4_data1, aes(depth, phosphate, color = block)) + geom_point() + facet_wrap( ~ block)

ggcorr(prob4_data1)

# because I can't find a correlation plot that separates by factor, I will create 8 separate data sets
detach(prob4_data1)
# prob4_data1.1 <- filter(prob4_data1, block == 1)
# prob4_data1.4 <- filter(prob4_data1, block == 4)
# prob4_data1.5 <- filter(prob4_data1, block == 5)
# prob4_data1.6 <- filter(prob4_data1, block == 6)
# prob4_data1.7 <- filter(prob4_data1, block == 7)
# prob4_data1.10 <- filter(prob4_data1, block == 10)
# prob4_data1.11 <- filter(prob4_data1, block == 11)
# prob4_data1.13 <- filter(prob4_data1, block == 13)

pairs(prob4_data1)

```

All 3 are numeric variables, but I converted block to a factor variable since it will be easier to see them as factors.

The lesosn in this dataset when considering the summary statistics and plots is that this data is meaningless and when seperated by block shows nice pictures.
