---
title: "HW5 STAT 5014"
author: "Shane Bookhultz"
date: "October 4, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

library(data.table)
library(stringr)
library(knitr)
library(dplyr)
library(downloader)
library(ggplot2)
library(fiftystater)

```


## Problem 3

What goes into making a good figure?
*Easy to comprehend
*Contains important information
*Makes a meaningful connection
*Able to be understood from all backgrounds

## Problem 4


a. This function successes makes a proportion of success from a vector of 1's and 0's and converts it to a proportion.

```{r echo = F}

Successes <- function(vec1) {
  # This is a function that takes in a vector and a number that they want to be the success of and returns the proportion of successes in the vector
  # vec1 is the vector it checks in
  # numsuc is the number that we want to see how many of them there are
  temp <- length(which(vec1 == 1))
  
  return (temp/length(vec1))
}

```
b. Create a matrix to simulate 10 flips of a coin with varying degrees of "fairness"

```{r echo = F}

set.seed(22495)
Flips10bino <- matrix(rbinom(10, 1, prob = (30:40)/100), nrow = 10, ncol = 10)

```

c. Use the function in conjunction with apply to compute successes by column and then row

```{r echo = F}

colprob <- apply(Flips10bino, 2, FUN = Successes)
rowprob <- apply(Flips10bino, 1, Successes)

```

What I notice here is that the column probabilities are all the same decimal value, but the row probabilities are all either 1's or 0's. 

d. Fix the matrix by creating a function whose input is a probability and output is a vector whose elements are the outcomes of 10 flips of a coin. 

```{r echo = F}

# not sure how to do this one

Flips10outcomes <- function(probability) {

  return (vec2)
}

```

## Problem 5 

```{r echo = F}

StarchData <- read.table("http://www2.isye.gatech.edu/~jeffwu/book/data/starch.dat", header = T, sep = "", stringsAsFactors = F)


kable(StarchData)

MeanStrengthCA <- mean(StarchData[,][2])

StrengthMean <- aggregate(strength~starch, StarchData, FUN = mean)
ThickMean <- aggregate(thickness~starch, StarchData, FUN = mean)

BetterViewMean <- inner_join(StrengthMean, ThickMean, by = "starch")

colnames(BetterViewMean) <- c("Starch", "StrengthMean", "ThicknessMean")

kable(BetterViewMean)

StarchData[,1] <- as.factor(StarchData[,1])

pairs(StarchData)

StarchModel <- lm(StarchData$strength ~ StarchData$starch + StarchData$thickness)

summary(StarchModel)
```

## Problem 6

```{r echo = F}

#we are grabbing a SQL set from here
#http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip

#download the files, looks like it is a .zip
download("http://www.farinspace.com/wp-content/uploads/us_cities_and_states.zip",dest="us_cities_states.zip")
unzip("us_cities_states.zip")
    
#read in data, looks like sql dump, blah
states <- fread(input = "./us_cities_and_states/states.sql",sep = "'", sep2 = ",", header = F, select = c(2,4))
### YOU do the CITIES  
### I suggest the cities_extended.sql may have everything you need
### can you figure out how to limit this to the 50?
# Why did you skip the 4 other states?
cities <- fread(input = "./us_cities_and_states/cities_extended.sql", sep = "'", sep2 = ",", header = F, select = c(2,4))

colnames(states) <- c("StateName", "StateAbbv")
colnames(cities) <- c("City", "StateAbbv")

citystate <- inner_join(cities, states, by = "StateAbbv")

statesum = table(citystate$StateName)
Numcitystate = as.data.frame(statesum)

colnames(Numcitystate) <- c("StateName", "Freq")

kable(Numcitystate)

# c. 

OccurLetter <- function (letter, state_name){
  # A function that counts the number of occurances of a letter in a string (state_name)
  # Output is a number with the count for that one letter
  # letter is a character
  # state_name is a string
  return (str_count(state_name, letter))
}

# loop through states

letter_count <- data.frame(matrix(NA,nrow = 51, ncol = 26))

for(k in 1:nrow(states)){
  for(j in 1:26) {
   letter_count[k,j] <- OccurLetter(letters[j], states[k,1]) 
  }
}

colnames(letter_count) <- letters

rownames(letter_count) <- states$StateName

# I don't know how to use the apply function with this

####### Plotting map

#https://cran.r-project.org/web/packages/fiftystater/vignettes/fiftystater.html

    
data("fifty_states") # this line is optional due to lazy data loading
crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
# map_id creates the aesthetic mapping to the state name column in your data
p <- ggplot(crimes, aes(map_id = state)) + 
  # map points to the fifty_states shape data
  geom_map(aes(fill = Assault), map = fifty_states) + 
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
    
p
#ggsave(plot = p, file = "HW5_Problem6_Plot_Settlage.pdf")

```

